Estimating dynamic structural equation models in R with the brms package

### 2025-01-28 index.qmd
Update & lint

p1 <- epred_average |> 
  ggplot(aes(hlag, .epred)) +
  geom_ribbon(
    aes(ymin = .lower, ymax = .upper),
    fill = "dodgerblue2", alpha = 0.2
  ) +
  geom_line(col = "dodgerblue2") +
  geom_point(
    data = dat |> 
      drop_na(happy),
    aes(y = happy),
    shape = 1, alpha = 0.5, size = 0.5,
    position = position_jitter(.01, .01)
  )

@@

fit1_fixef_ellipse <- ellipse(
  cor(as.data.frame(fit1)[,1:2])[1,2],
  scale = fit1_fixef_ellipse$sd, centre = fit1_fixef_ellipse$mean
) |> as_tibble()

p3 <- fit1_coef |> 
  left_join(fit1_coef_lm) |> 
  ggplot(aes(x = `estimate_(Intercept)`, y = estimate_hlag)) +
  scale_x_continuous(
    "Intercept",
    expand = expansion(-0.1)
  ) +
  scale_y_continuous(
    "Slope",
    expand = expansion(-0.2)
  ) +
  # geom_segment(
  #   aes(xend = Estimate.b0, yend = Estimate.b1),
  #   linewidth = 0.1, alpha = 0.5
  # ) +
  geom_point(
    aes(`estimate_(Intercept)`, estimate_hlag),
    col = "gray80", size = 0.75
  ) +
  geom_point(
    aes(x = Estimate.b0, y = Estimate.b1), 
    col = "black", size = 1
  ) +
  geom_point(
    data = fit1_fixef,
    aes(
      x=mean_b_b0, 
      y=mean_b_b1
    ), 
    col = "dodgerblue2", size = 2
  ) +
  geom_path(
    data = fit1_fixef_ellipse,
    aes(x, y),
    col = "dodgerblue", linewidth = 0.75
  ) +
  theme(aspect.ratio = 1)

@@

dat <- dat |> 
  pivot_longer(sm:happy, names_to = "outcome", values_to = "y") |> 
  mutate(
    Ih = if_else(outcome == "happy", 1, 0) |> as.integer(),
    Ism = if_else(outcome == "sm", 1, 0) |> as.integer(), 
    .before = y
  ) |> 
  left_join(dat)

dat2 <- dat |> 
  select(class, person, time, Ih, Ism, sm, happy, hlag, y)

@@

epred_average <- dat |> 
  data_grid(
    sm = seq_range(sm, n = 9),
    hlag = mean(hlag, na.rm = TRUE),
    Ih = 1, Ism = 0
  ) |> 
  add_epred_draws(fit2, re_formula = NA) |> 
  mean_qi()

epred_person <- dat |> 
  filter(person %in% tmp$person) |> 
  group_by(person) |> 
  data_grid(
    sm = seq_range(sm, n = 9),
    hlag = mean(hlag, na.rm = TRUE),
    Ih = 1, Ism = 0
  ) |> 
  add_epred_draws(fit2) |> 
  mean_qi()

p1 <- epred_average |> 
  ggplot(aes(sm, .epred)) +
  geom_ribbon(
    aes(ymin = .lower, ymax = .upper),
    fill = "grey80"
  ) +
  geom_line() +
  geom_point(
    data = dat |> 
      filter(Ih == 1) |> 
      drop_na(happy),
    aes(y = happy),
    shape = 1, alpha = 0.5,
    position = position_jitter(.01, .01)
  )

p2 <- epred_person |>   
  ggplot(aes(sm, .epred)) +
  geom_ribbon(
    aes(ymin = .lower, ymax = .upper),
    fill = "grey80"
  ) +
  geom_line() +
  geom_point(
    data = dat |> 
      filter(person %in% tmp$person, Ih == 1) |> 
      drop_na(happy),
    aes(y = happy),
    shape = 1, alpha = 0.5,
    position = position_jitter(.01, .01)
  ) +
  facet_wrap("person")

@@

fit3 <- brm(
  bform,
  data = dat,
  control = list(adapt_delta = 0.99),
  file = "cache/brm-3"
)
```

>>>>

p1 <- epred_average |>
  ggplot(aes(hlag, .epred)) +
  geom_ribbon(
    aes(ymin = .lower, ymax = .upper),
    fill = "dodgerblue2", alpha = 0.2
  ) +
  geom_line(col = "dodgerblue2") +
  geom_point(
    data = dat |>
      drop_na(happy),
    aes(y = happy),
    shape = 1, alpha = 0.5, size = 0.5,
    position = position_jitter(.05, .05)
  )

@@

fit1_fixef_ellipse <- ellipse(
  x = cor(as.data.frame(fit1)[,1:2])[1,2],
  scale = fit1_fixef_ellipse$sd,
  centre = fit1_fixef_ellipse$mean
) |> as_tibble()

p3 <- fit1_coef |>
  left_join(fit1_coef_lm) |>
  ggplot(aes(x = `estimate_(Intercept)`, y = estimate_hlag)) +
  scale_x_continuous(
    "Intercept",
    expand = expansion(-0.1)
  ) +
  scale_y_continuous(
    "Slope",
    expand = expansion(-0.2)
  ) +
  geom_point(
    aes(`estimate_(Intercept)`, estimate_hlag),
    col = "gray80", size = 0.75
  ) +
  geom_point(
    aes(x = Estimate.b0, y = Estimate.b1),
    col = "black", size = 1
  ) +
  geom_point(
    data = fit1_fixef,
    aes(
      x=mean_b_b0,
      y=mean_b_b1
    ),
    col = "dodgerblue2", size = 2
  ) +
  geom_path(
    data = fit1_fixef_ellipse,
    aes(x, y),
    col = "dodgerblue", linewidth = 0.75
  ) +
  theme(aspect.ratio = 1)

@@

dat2 <- dat |>
  pivot_longer(sm:happy, names_to = "outcome", values_to = "y") |>
  mutate(
    Ih = if_else(outcome == "happy", 1, 0) |> as.integer(),
    Ism = if_else(outcome == "sm", 1, 0) |> as.integer(),
    .before = y
  ) |>
  left_join(dat)

dat2 <- dat2 |>
  select(class, person, time, Ih, Ism, sm, happy, hlag, y)

@@

epred_average <- dat2 |>
  data_grid(
    sm = seq_range(sm, n = 9),
    hlag = mean(hlag, na.rm = TRUE),
    Ih = 1, Ism = 0
  ) |>
  add_epred_draws(fit2, re_formula = NA) |>
  mean_qi()

epred_person <- dat2 |>
  filter(person %in% tmp$person) |>
  group_by(person) |>
  data_grid(
    sm = seq_range(sm, n = 9),
    hlag = mean(hlag, na.rm = TRUE),
    Ih = 1, Ism = 0
  ) |>
  add_epred_draws(fit2) |>
  mean_qi()

p1 <- epred_average |>
  ggplot(aes(sm, .epred)) +
  geom_ribbon(
    aes(ymin = .lower, ymax = .upper),
    fill = "grey80"
  ) +
  geom_line() +
  geom_point(
    data = dat2 |>
      filter(Ih == 1) |>
      drop_na(happy),
    aes(y = happy),
    shape = 1, alpha = 0.5,
    position = position_jitter(.01, .01)
  )

p2 <- epred_person |>
  ggplot(aes(sm, .epred)) +
  geom_ribbon(
    aes(ymin = .lower, ymax = .upper),
    fill = "grey80"
  ) +
  geom_line() +
  geom_point(
    data = dat2 |>
      filter(person %in% tmp$person, Ih == 1) |>
      drop_na(happy),
    aes(y = happy),
    shape = 1, alpha = 0.5,
    position = position_jitter(.01, .01)
  ) +
  facet_wrap("person")

@@

fit3 <- brm(
  bform,
  data = dat2,
  control = list(adapt_delta = 0.99),
  file = "cache/brm-3"
)
```
